---
title: "Lab_Introduction_to_R"
author: "Singh, Gurjeet"
date: "April 7, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Basic Commands

The following command instructs R to join together the numbers 1, 3, 2, and 5, and to save them as a vector named x. When we type x, it vector gives us back the vector.

```{r L1}
x <- c(1,3,2,5)
x

```

Note that the > is not part of the command; rather, it is printed by R to indicate that it is ready for another command to be entered. We can also save things using = rather than <-:

```{r L1.1}
x = c(1, 6, 2)
x

y = c(1,4,3)
y
```

## 2. Graphics

### 2.1. plot()

The plot() function is the primary way to plot data in R. For instance, plot(x,y) produces a v of the numbers in x versus the numbers in y. There are many additional options that can be passed in to the plot() function. For example, passing in the argument xlab will result in a label on the x-axis.

```{r L2}
x=rnorm (100)
y=rnorm (100)
plot(x,y)
plot(x,y,xlab=" this is the x-axis",ylab=" this is the y-axis",
main=" Plot of X vs Y")
```

### 2.2. pdf() and jpeg ()

We will often want to save the output of an R plot. The command that we use to do this will depend on the file type that we would like to create. For instance, to create a pdf, we use the pdf() function, and to create a jpeg, we use the jpeg() function.

The function dev.off() indicates to R that we are done creating the plot.

```{r L2.1}
# Creating PDF
pdf (" Figure.pdf ")
plot(x,y,col =" green ")
dev.off ()

#Creating JPEG
jpeg (" Figure.jpeg ")
plot(x,y,col =" green ")
dev.off ()
```

### 2.3. seq()

The function seq() can be used to create a sequence of numbers. For instance, seq(a,b) makes a vector of integers between a and b. There are many other options: for instance, seq(0,1,length=10) makes a sequence of 10 numbers that are equally spaced between 0 and 1. Typing 3:11 is a shorthand for seq(3,11) for integer arguments.

```{r L2.2}
 x=seq (1 ,10)
 x

 x=1:10
 x

 x=seq(-pi ,pi ,length =50)
 x
 
 seq(0,1,length=10)
```

### 2.4. contour()

We will now create some more sophisticated plots. The contour() function produces a contour plot in order to represent three-dimensional data; it is like a topographical map. It takes three arguments:

1. A vector of the x values (the first dimension),
2. A vector of the y values (the second dimension), and
3. A matrix whose elements correspond to the z value (the third dimension) for each pair of (x,y) coordinates.

As with the plot() function, there are many other inputs that can be used to fine-tune the output of the contour() function. To learn more about these, take a look at the help file by typing ?contour.

```{r L2.3}
y=x
f=outer(x,y,function (x,y)cos(y)/(1+x^2))
contour(x,y,f)
contour(x,y,f,nlevels =45, add=T)
fa=(f-t(f))/2
contour (x,y,fa,nlevels =15)
```

### 2.5. image() and persp()

The image() function works the same way as contour(), except that it produces a color-coded plot whose colors depend on the z value. This is known as a heatmap, and is sometimes used to plot temperature in weather forecasts. Alternatively, persp() can be used to produce a three-dimensional plot. The arguments theta and phi control the angles at which the plot is viewed.

```{r L2.4}
image(x,y,fa)
persp(x,y,fa)
persp(x,y,fa ,theta =30)
persp(x,y,fa ,theta =30, phi =20)
persp(x,y,fa ,theta =30, phi =70)
persp(x,y,fa ,theta =30, phi =40)
```


## 3. Indexing Data

### 3.1. matrix()

We often wish to examine part of a set of data. Suppose that our data is
stored in the matrix A.

```{r L3}
A=matrix (1:16 ,4 ,4)
A
```

Then, typing will select the element corresponding to the second row and the third column. The first number after the open-bracket symbol [ always refers to the row, and the second number always refers to the column. We can also select multiple rows and columns at a time, by providing vectors as the indices.

```{r L3.1}
A[2,3]

A[c(1,3) ,c(2,4) ]

A[1:3 ,2:4]

A[1:2 ,]

A[ ,1:2]

```

The last two examples include either no index for the columns or no index for the rows. These indicate that R should include all columns or all rows, respectively. R treats a single row or column of a matrix as a vector.

```{r L3.2}

A[1,]

```


The use of a negative sign - in the index tells R to keep all rows or columns except those indicated in the index.

```{r L3.3}

A[-c(1,3) ,]

A[-c(1,3) ,-c(1,3,4)]
```

The dim() function outputs the number of rows followed by the number of columns of a given matrix.

```{r L3.4}
dim(A)
```


## 4. Loading Data

### 4.1. data()

We begin by loading in the Auto data set. The following command will load the Auto.data file into R and store it as an object called Auto, in a format referred to as a data frame.

```{r L4}
#install.packages("ISLR")
library(ISLR)
data(Auto)
head(Auto)
#fix(Auto)
```

### 4.2. names()

Once the data are loaded correctly, we can use names() to check the variable names.

```{r L4.1}

names(Auto)
```

## 5. Additional Graphical and Numerical Summaries

### 5.1. plot() and attach()

To refer to a variable, we must type the data set and the variable name joined with a $ symbol. Alternatively, we can use the attach() function in order to tell R to make the variables in this data frame available by name

```{r L5.1}

plot(Auto$cylinders , Auto$mpg )
attach (Auto)
plot(cylinders , mpg)

```

### 5.2. as.factor()

The cylinders variable is stored as a numeric vector, so R has treated it as quantitative. However, since there are only a small number of possible values for cylinders, one may prefer to treat it as a qualitative variable. The as.factor() function converts quantitative variables into qualitative variables.

```{r L5.2}
cylinders =as.factor (cylinders )
```

### 5.3. plot() with options

If the variable plotted on the x-axis is categorial, then boxplots will boxplot automatically be produced by the plot() function. As usual, a number of options can be specified in order to customize the plots.

```{r L5.3}

plot(cylinders , mpg)
plot(cylinders , mpg , col ="red ")
plot(cylinders , mpg , col ="red", 
      varwidth =T)
plot(cylinders , mpg , col ="red", 
      varwidth =T,horizontal =T)
plot(cylinders , mpg , col ="red", 
      varwidth =T, xlab=" cylinders ", ylab ="MPG ")

```

### 5.4. Histogram (hist())

The hist() function can be used to plot a histogram. Note that col equals to 2 histogram has the same effect as col equals to "red".

```{r L5.4}

hist(mpg)
hist(mpg ,col =2)
hist(mpg ,col =2, breaks =15)

```

### 5.5. pairs()

The pairs() function creates a scatterplot matrix i.e. a scatterplot for every scatterplot pair of variables for any given data set. We can also produce scatterplots matrix for just a subset of the variables.

```{r L5.5}
pairs(Auto)
pairs(~ mpg  + displacement 
      + horsepower 
      + weight 
      + acceleration
      , Auto)

plot(Auto$displacement, Auto$mpg)
plot(Auto$horsepower, Auto$mpg)
plot(Auto$weight, Auto$mpg)
plot(Auto$acceleration, Auto$mpg)
```

### 5.6. identify()

In conjunction with the plot() function, identify() provides a useful interactive method for identifying the value for a particular variable for points on a plot. We pass in three arguments to identify(), the x-axis variable, the y-axis variable, and the variable whose values we would like to see printed for each point. Then clicking on a given point in the plot will cause R to print the value of the variable of interest. Right-clicking on the plot will exit the identify() function (control-click on a Mac). The numbers printed under the identify() function correspond to the rows for the selected points.

```{r L5.6}
plot(horsepower ,mpg)
identify (horsepower ,mpg ,name)
```


### 5.7. Summary()

The summary() function produces a numerical summary of each variable in a particular data set.

```{r L5.7}
summary (Auto)
```

For qualitative variables such as name, R will list the number of observations that fall in each category. We can also produce a summary of just a single variable.

```{r L5.8}
summary (mpg)
```










